---
title: "Analysis of Raw Data"
output: html_notebook
---

Set up packages needed for analysis
```{r message=FALSE, warning=FALSE}
packages.used = c("readxl", "xlsx", "tidyr", "ggplot2", "stringr", "lubridate", "stringi", "dplyr","uuid","scales", "purrr", "readr", "clipr")

# check packages that need to be installed.
packages.needed=setdiff(packages.used, 
                        intersect(installed.packages()[,1], 
                                  packages.used))
# install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE)
}

output <- lapply(packages.used, require, character.only = TRUE)

all(output) # true if all code libraries are correctly loaded
```

Specify the raw data file
```{r}
file_raw_data <- file.path("raw_annual_data.xlsx")

raw_data_sheets <- excel_sheets(path = file_raw_data)
```

Import the data from the raw data file sheet by sheet
```{r message=FALSE, warning=FALSE}
df <- raw_data_sheets %>%
  map_df(~ read_excel(path = file_raw_data, sheet = .x, 
                      skip = 1,
                      col_names = c("Ticker", "FY06", "FY07", "FY08", "FY09", "FY10", "FY11", "FY12", "FY13", "FY14", "FY15", "FY16"),
                      col_types = c("text","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric"),
                      na = c("#N/A N/A"),
                      ) %>% mutate(sheet = .x))
```

Transform the data so that we have the FY in a single column (tidy format).
```{r}
eps <- df %>%
  gather(key = "fiscal_year", value = "value", starts_with("FY")) %>%
  rename(ticker = Ticker,
         value_type = sheet) %>%
  mutate(fiscal_year = as.integer(str_replace_all(fiscal_year, "FY","20")),
         ticker = str_extract(ticker, "^([A-Z^\\w]*)")) %>%
  drop_na(value)

str(eps)
```

Import and join data on the sector specification of each S&P 500 component.
```{r message=FALSE, warning=FALSE}
ind <- read_csv(file = "constituents_csv.csv")
colnames(ind) <- colnames(ind) %>% str_to_lower()

tj <- left_join(eps, ind, by = c("ticker" = "symbol"))
```

Filter down the guidance and eps actuals we are interested in.  Then drop empty cells and compute the variance and forecast error for each observation.
```{r}
eps <- tj %>%
  filter(value_type %in% c("Annual_EPS GAAP_Guid", "Annual_EPS GAAP_Act (Dil)")) %>%
  mutate(value_type = 
           case_when(
             value_type == "Annual_EPS GAAP_Guid" ~ "eps_fcst",
             value_type == "Annual_EPS GAAP_Act (Dil)" ~ "eps_act"
           )) %>%
  unique() %>%
  group_by(ticker, fiscal_year, value_type) %>%
  spread(value_type, value) %>%
  drop_na(eps_fcst, eps_act) %>%
  mutate(eps_var_pct = (eps_fcst - eps_act) / eps_act,
         fcst_err = abs(eps_var_pct))
```

Compute the average forecast error for each company and then rank each company by average forecast error relative to other companies within its sector.
```{r}
fcst_rank <- eps %>% 
  group_by(sector) %>%
  mutate(avg_sector_fcst_err = mean(fcst_err)) %>%
  ungroup() %>%
  group_by(sector, ticker, name) %>%
  summarise(avg_comp_fcst_err = mean(fcst_err)) %>%
  ungroup() %>%
  group_by(sector) %>%
  mutate(fcst_rank_sector = min_rank(avg_comp_fcst_err)) %>%
  arrange(sector, desc(-fcst_rank_sector))

fcst_rank
```

